---
title: "My Basic Movement Report"
subtitle: "created with pkg stmove" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
params:
  df: NA
  stats: NA
  na.approx: NA
  seas: NA
  construct: NA
  proj4: NA
  wavelet: NA
---

```{r setup, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE
)

library(patchwork)

construct <- (!is.null(params$construct))
klocoh <- "klocoh" %in% tolower(params$construct)
akde <- "akde" %in% tolower(params$construct)
rolling <- ("rolling" %in% params$stats)
diurnal <- ("diurnal" %in% params$stats)
lunar <- ("lunar" %in% params$stats)
seasonal <- ("seasonal" %in% params$stats)
```

# Visualize my trajectory

```{r}
plot(params$df$x, params$df$y, type = "p")
```

# Distributions:

```{r, fig.height = 4}
ss <- ss_dist(x = params$df)
ta <- ta_dist(x = params$df)
```

# Basic Statistical Summary

`r if (rolling) '## Fine Scale Sliding Window'`

```{r eval = rolling, messages = F, fig.height = 4}
roll <<- rolling_stats(params$df, na.approx = params$na.approx)

p1 <- ggplot(roll, aes(date, y = mean_dist)) +
  geom_point(na.rm = T) + geom_smooth(na.rm = T)

p2 <- ggplot(roll, aes(date, y = mean_ang)) +
  geom_point(na.rm = T) + geom_smooth(na.rm = T)


p1 + p2 + plot_annotation("Mean Step Size & Turning Angle on a Rolling Interval")

ggplot(roll, aes(date)) +
  geom_smooth(aes(y = acf_ang, color = "Relative Angle"), na.rm = T) +
  geom_smooth(aes(y = acf_dist, color = "Step Size"), na.rm = T) +
  geom_smooth(aes(y = ccf, color = "Cross Correlation"), na.rm = T) +
  ylab("Correlation") +
  ggtitle("Lag 1 Auto and Cross Correlations in Turning Angle & Step Size",
    subtitle = "Rolling Window"
  )
```

`r if (diurnal) '## Diurnal Cycle'`

```{r eval = diurnal, message = FALSE}
diurnal_stats <<- interval_stats(params$df)

# We need to make some decisions about how to plot these things...
plots <- purrr::map(
  names(diurnal_stats)[c(2, 5, 4, 7, 8)],
  function(x){
    y <- sym(x)
  p <- ggplot(
    diurnal_stats,
    aes(interval_start, y = !!y,  color = ifelse(hour(interval_start) == 0,
        "0:00 - 12:00",
        "12:00 - 24:00"
      )
    )
  ) + geom_point(na.rm = T, size = .3) + geom_smooth(na.rm = T)  +
    xlab(NULL) +
theme(axis.text.x = element_text(angle = 25, vjust = .5))
  
  if(x == "acf_ang"){
    return(p + scale_color_discrete("Time of Day") + 
             theme(legend.position = "bottom"))
  } else{
     return(p + guides(color = "none"))
  }
  }
    )

{plots[[1]] + plots[[2]]} - {plots[[3]] + plots[[4]]+ plots[[5]]} + plot_layout(nrow = 2) +
  plot_annotation(title = "Diurnal Interval Statistics")

```

`r if (lunar) '## Lunar Cycle'`

```{r eval = lunar}
lunar_stats <<- interval_stats(params$df, type = "lunar")
```

`r if (seasonal) '## Seasonal Cycle'`

```{r eval = seasonal}
seasonal_stats <<- interval_stats(params$df, type = "seasonal", seas = params$seas)
```

`r if (params$wavelet) '# Visualizing Periodicity'`

```{r eval= params$wavelet}

```

`r if (construct) '# Space Use Constructions'`

`r if (klocoh) '## A k-LoCoh hull set'`

```{r, eval = klocoh}
locoh <<- construct(params$df, type = "klocoh", proj4 = params$proj4)
```

`r if (akde) '## An autocorrelated kernel density estimation'`
```{r, eval = akde}
akde <<- construct(params$df, type = "akde", proj4 = params$proj4)
```

\newpage
# Suggested Next Steps

This report provides a first-pass look at your trajectories and patterns you see here
may be further investigated in any number of ways. Below we suggest some common 
next questions and steps and associated packages that could help you begin to address them. 
Also see suggestions from Seidel et al. 2019, Appendix C of Seidel et al. 2018,
or tutorials from the 2018 Movement Ecology in R workshop hosted [here](https://www.danaseidel.com/MovEco-R-Workshop/).

- Habitat Selection Analysis
    - the `sf` package offers ...
    - the `raster` package offers ...
    - the `lme4` package is useful for fitting generalized linear models with fixed or random effects
- Hidden Markov Models
    - `momentuHMM`
    - `moveHmm`
- Dynamic Interaction between individuals?
    - `wildlifeDI`

# Constructing Home Ranges
=======
Also consider the numerous analyses made available by the extentive "adehabitat" packages:
`adehabitatLT`, `adehabitatHR`, `adehabitatMA`, `adehabitatHS` and their indepth vignettes. 
