---
title: "Movement Report for `r params$df$id[1]`"
subtitle: "created with pkg stmove" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
geometry: margin=2cm
params:
  df: NA
  stats: NA
  seas: NA
  construct: NA
  proj4: NA
  wavelet: NA
---

```{r setup, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(stmove)
library(patchwork)

construct <- (!is.null(params$construct))
klocoh <- "klocoh" %in% tolower(params$construct)
akde <- "akde" %in% tolower(params$construct)
rolling <- ("rolling" %in% params$stats)
diurnal <- ("diurnal" %in% params$stats)
lunar <- ("lunar" %in% params$stats)
seasonal <- ("seasonal" %in% params$stats)
wavelet <- (!is.null(params$wavelet))
```

# Visualize the trajectory

```{r}
ggplot(params$df, aes(x, y)) + geom_point() + labs(x = "Easting", y = "Northing") +
  theme_minimal()
```

# Step Length & Turning Angle Distributions:

```{r, fig.height = 4}
ss <- ss_dist(x = params$df)
```

```{r, fig.height = 4}
ta <- ta_dist(x = params$df)
```

`r if (any(c(rolling, diurnal, lunar, seasonal))) '# Basic Statistical Summary'`

`r if (rolling) '## Fine Scale Sliding Window'`

```{r eval = rolling, messages = F, fig.height = 4}
roll <<- rolling_stats(params$df)

ggplot(roll, aes(date, y = mean_dist)) +
  geom_path(na.rm = T) +
  facet_wrap(~ paste(lubridate::month(date, label = TRUE), lubridate::year(date)),
    scales =
      "free_x"
  ) +
  labs(
    title = "Mean Step Size on a Rolling interval",
    y = "Mean Step Size (m)"
  ) + theme_minimal() +
  scale_x_datetime("Day of the month", date_labels = "%d")


ggplot(roll, aes(date, y = mean_ang)) +
  geom_path(na.rm = T) +
  facet_wrap(~ paste(lubridate::month(date, label = TRUE), lubridate::year(date)),
    scales =
      "free_x"
  ) +
  labs(
    title = "Mean Turning Angle on a Rolling interval",
    y = "Mean Angle (radians)"
  ) + theme_minimal() +
  scale_x_datetime("Day of the month", date_labels = "%d")

suppressMessages(ggplot(roll, aes(date)) +
  geom_smooth(aes(y = acf_ang, color = "Relative Angle"), na.rm = T) +
  geom_smooth(aes(y = acf_dist, color = "Step Size"), na.rm = T) +
  geom_smooth(aes(y = ccf, color = "Cross Correlation"), na.rm = T) +
  ylab("Correlation") +
  ggtitle("Lag 1 Auto and Cross Correlations in Turning Angle & Step Size",
    subtitle = "Rolling Window"
  )) + theme_minimal()
```

`r if (diurnal) '## Diurnal Cycle'`

```{r eval = diurnal, message = FALSE}
diurnal_stats <<- interval_stats(params$df)

# We need to make some decisions about how to plot these things...
plots <- purrr::map(
  names(diurnal_stats)[c(3, 6, 5, 8, 9)],
  function(x) {
    y <- sym(x)
    p <- ggplot(
      diurnal_stats,
      aes(interval_start, y = !!y, color = TOD)
    ) + geom_point(na.rm = T, size = .3) +
      geom_smooth(na.rm = T) +
      xlab(NULL) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 25, vjust = .5))

    if (x == "acf_ang") {
      return(p + scale_color_discrete("Time of Day") +
        theme(legend.position = "bottom"))
    } else {
      return(p + guides(color = "none"))
    }
  }
)

suppressMessages({
  plots[[1]] + plots[[2]]
} - {
  plots[[3]] + plots[[4]] + plots[[5]]
} + plot_layout(nrow = 2) +
  plot_annotation(title = "Diurnal Interval Statistics"))
```

`r if (lunar) '## Lunar Cycle'`

```{r eval = lunar}
lunar_stats <<- interval_stats(params$df, type = "lunar")

# We need to make some decisions about how to plot these things...
plots <- purrr::map(
  names(lunar_stats)[c(3, 6, 5, 8, 9)],
  function(x) {
    y <- sym(x)
    p <- ggplot(
      lunar_stats,
      aes(x = as.numeric(interval_start), y = !!y, color = phase)
    ) +
      geom_step(aes(group = 1)) +
      # scale_x_continuous(breaks = 1:nlevels(lunar_stats$interval_start),
      #                   labels = levels(lunar_stats$interval_start)) +
      xlab(NULL) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 25, vjust = .5))

    if (x == "acf_ang") {
      return(p + scale_color_discrete("Lunar Phase") +
        theme(legend.position = "bottom"))
    } else {
      return(p + guides(color = "none"))
    }
  }
)

{
  plots[[1]] + plots[[2]]
} - {
  plots[[3]] + plots[[4]] + plots[[5]]
} + plot_layout(nrow = 2) +
  plot_annotation(title = "Lunar Interval Statistics")
```

`r if (seasonal) '## Seasonal Cycle'`

```{r eval = seasonal}
seasonal_stats <<- interval_stats(params$df, type = "seasonal", seas = params$seas)

p1 <- seasonal_stats %>%
  # tidyr::gather("stat", "estimate", -1) %>%
  ggplot(., aes(x = as.numeric(interval_start), y = mean_dist)) +
  geom_col() +
  geom_errorbar(aes(
    ymin = mean_dist,
    ymax = mean_dist + sd_dist
  )) + labs(x = "", y = "Average Step Size (+ 1 SD)")
p2 <- seasonal_stats %>%
  ggplot(., aes(x = as.numeric(interval_start), y = mean_ang)) +
  geom_col() +
  labs(x = "Season", y = "Mean Turning Angle")

p3 <- seasonal_stats %>%
  select(-mean_dist, -sd_dist, -mean_ang, -sd_ang) %>%
  tidyr::gather("stat", "estimate", -1) %>%
  ggplot(., aes(
    x = as.factor(as.numeric(interval_start)), y = estimate,
    group = stat, color = stat
  )) +
  geom_point() +
  geom_line() +
  lims(y = c(-1, 1)) +
  labs(x = "")

p1 + p2 + p3 +
  plot_annotation(
    title = "Interval Statistics Across Seasons",
    theme = theme(plot.title = element_text(hjust = .5))
  ) &
  theme_minimal()
```

`r if (wavelet) '# Visualizing Periodicity'`

```{r eval = wavelet}
wave <<- wavelet(params$df, stats = params$wavelet, useRaster = TRUE)
```

`r if (construct) '# Space Use Constructions'`

`r if (klocoh) '## A k-LoCoh hull set'`

```{r, eval = klocoh}
locoh <<- construct(params$df, type = "klocoh", proj4 = params$proj4)
```

`r if (akde) '## An autocorrelated kernel density estimation'`
```{r, eval = akde}
AKDE <<- construct(params$df, type = "akde", proj4 = params$proj4)
```

\newpage
# Suggested Next Steps

This report provides a first-pass look at your trajectories and patterns you see here
may be further investigated in any number of ways. Below we suggest some common 
next questions and steps and associated packages that could help you begin to address them. 
Also see suggestions from Seidel et al. 2019, Appendix C of Seidel et al. 2018,
or tutorials from the 2018 Movement Ecology in R workshop hosted [here](https://www.danaseidel.com/MovEco-R-Workshop/).

- Habitat Selection Analysis
    - the `sf` package offers tools for ready manipulation of spatial vector data. 
    - the `raster` and `velox` packages offer tools for raster data manipulation. 
    - the `lme4` package is useful for fitting generalized linear models with fixed or random effects
- Hidden Markov Models can be fit or Behavioral Change Point Analysis done with tools from:
    - `momentuHMM`
    - `moveHmm`
    - `bcpa`
- Dynamic Interaction between individuals may be explored use
    - `wildlifeDI`
    
Also consider the numerous analyses made available by the extentive "adehabitat" packages:
`adehabitatLT`, `adehabitatHR`, `adehabitatMA`, `adehabitatHS` and their indepth vignettes. 
